Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-01-30 numerical-analysis.tex
Adding shock-analysis section, including one citation and one newcommand \Mach.

>>>>

Identification of shocks and their pre- and post-shock conditions can be
accomplished through a combination of either temperature or velocity jumps with
dimensionally split or unsplit search methods.  The primary method used in
\enzo~is the dimensionally unsplit temperature jump method described in detail
in \citet{2008ApJ...689.1063S}.  We briefly outline the method below.

For every cell, we first determine whether is satisfies the following conditions
necessary to be flagged as a shock:
\begin{eqnarray}
\nabla \cdot \vec{v} < 0 \\
\nabla T \cdot \nabla S > 0\\
T_2 > T_1 \\
\rho_2 > \rho_1,
\end{eqnarray}
where $\vec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
density, and $S=T/\rho^{\gamma-1}$ is the entropy.  An optional temperature floor
may be chosen, useful for example in cosmological simulations without a radiation
background.

Once a cell is flagged, the local temperature gradient is calculated, which is 
then used to traverse cells parallel to the gradient to search for the first
pre- and post-shock cell that do not satisfy the above conditions.  If during
the search a cell satisfying the conditions is found to have a more convergent
flow, that cell is marked as the center, and the search is started again. Using the 
temperature values from each of these cells, the Mach number is then solved
using the Rankine-Hugoniot temperature jump conditions:
\begin{equation}
\frac{T_2}{T_1} = \frac{(5\Mach^2 - 1)(\Mach^2 + 3)}{16\Mach^2},
\end{equation}
where $T_2$ and $T_1$ are the post-shock~(downstream) and pre-shock~(upstream)
temperatures, respectively. $\Mach$ is the upstream Mach number.

Shock finding in the context of AMR is applied grid-by-grid.  If a search for
pre/post-shock cells goes outside the bounds of the grid ghost zones, the
search is stopped for that particular shocked cell. In most situations this is
adequate since the hydrodynamic shock is captured in fewer than the number of
ghost zones.  Shock finding can be run either upon data output, or at each step
in the evolution of an AMR level if the Mach and pre/post-shock quantities are
needed for additional physics modules.
